Eco-evolutionary dynamics of finite populations from first principles

### 2023-04-24 ind_files/06_nontechnical_discussion.tex
Add connections to Rice papers

>>>>

\textcolor{blue}{Rice has proposed a stochastic version of the Price equation in a series of papers~\citep{rice_stochastic_2008,rice_evolution_2009,rice_universal_2020} that follow the classic derivation by Price himself~\citep{rice_evolutionary_2004}. Like the original Price equation, these equations are formulated in a very general manner that relates the phenotypic change between two given populations. As such, they are more general than my equations, but, just like the original Price equation, are dynamically insufficient and cannot be phrased in the language of dynamical systems (to the best of my understanding). They are thus the true stochastic analog of the original Price equation, whereas my version (equation \eqref{nD_stochastic_Price}) is the analog of~\posscite{lion_theoretical_2018} version of the Price equation in a continuous time, dynamically sufficient setting. The two formulations are complementary --- Rice's equations (and the original Price equation) provide a very general description for partitioning phenotypic change between two time points, whereas my approach (and ~\posscite{lion_theoretical_2018} version of the Price equation) provides equations that are slightly less general, but are formulated in a continuous time, dynamically sufficient, predictive manner. Rice's derivations also treat fitness as fundamental, whereas I derive suitable notions of fitness and turnover from demographic first principles. As a consequence, the `extra' stochastic term corresponding to noise-induced selection that appears in my equations fundamentally emerges from the stochasticity of the underlying births and deaths of organisms and is thus of ecological/demographic origin, whereas the `extra' stochastic term in Rice's equations emerges from the stochasticity of fitness\footnote{defined here as the number of offspring produced --- no deaths involved~\citep{rice_stochastic_2008,rice_evolution_2009,rice_universal_2020}}alone and, to the best of my knowledge, does \emph{not} correspond to the same effect I identify in this thesis. Lastly, whereas Rice's equations are exact, my equations are derived using a system-size expansion (see part \ref{part_theory}) and thus are, strictly, speaking only an approximate statistical description of the dynamics (though these approximations are usually very good for even moderate values of $K$~\citep{black_stochastic_2012,cianci_wkb_2015})}
